---
title: "Exploratory_Analysis"
author: "Serena Zhang"
date: "4/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE, 
                      cache = TRUE)
library(tidyverse)
library(lattice)
library(DAAG)
library(tidyr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(grid)
```
## Reading Data
```{r}
art = read.csv("final_sothebys.csv", header=TRUE, na.strings=c("", NA))
#Save as a data frame
art_df <- as.data.frame(art)
#Dropping features with no interest (non-informative) in the analysis
drop <- c("X", "Unnamed..0", "provenance", "auction_house_id", "external_image_url", "literature","condition","auc_desc")
art_df <- art_df[, !(names(art_df) %in% drop)]
```
```{r}
df1 <- art_df %>% drop_na(hammer_price_bp_usd) 
df2 <- df1 %>%
  filter(abs(df1$hammer_price_bp_usd - 
               median(df1$hammer_price_bp_usd)) <=3*sd(df1$hammer_price_bp_usd)) 
#df1 <- filter(df1,!df1$loation %in% c("DOHA","HONG KONG"))
```
## Auction Information

```{r }
#ggplot(df2,aes(x=hammer_price_bp_usd))+geom_histogram(aes(y=..density..))+geom_density()
```

## Do Certain Lot Attributes Result in Higher Price?
### lot titles
#### What lots have higher price? 
```{r}
df_wordcloud <- df1[,c("lot_title","hammer_price_bp_usd")]
df_wordcloud <- arrange(df_wordcloud,desc(df_wordcloud$hammer_price_bp_usd))[1:500,]
library(wordcloud)
library(tm)
pal <- brewer.pal(9, "OrRd")
pal <- pal[-(1:3)]
wordcloud(df_wordcloud$lot_title, df_wordcloud$hammer_price_bp_usd, min.freq=500, scale=c(5, .5), random.order = FALSE, random.color = FALSE, colors= pal)
```
#### What are the top words in lot title that are related to higher price? 
```{r}
# collapse the lot_title column by word and count the frequency they appear in the titles.
temp <- paste(df1$lot_title, collapse=' ' )
temp <- tolower(temp)
temp <- gsub(" *\\b[[:alpha:]]{1}\\b *", " ", temp)
temp <- gsub('[[:punct:] ]+',' ',temp)
temp <- as.list(strsplit(temp, " "))
temp <- unlist(temp)[!(unlist(temp) %in% stopwords("english"))]
temp <- unlist(temp)[!(unlist(temp) %in% "na")]
word_count <- na.omit(as.data.frame(table(temp)))
word_count <- arrange(word_count,desc(word_count$Freq))[1:300,]

# visualize word frequencies
pal <- brewer.pal(9, "Dark2")
wordcloud(word_count$temp, word_count$Freq, min.freq =20, scale=c(5, .5), random.order = FALSE, random.color = FALSE, colors= pal)
```
#### Looking at Untitle, what price ranges are they in? Is it correlated?
```{r}
library(vcd)
df1 <- df1 %>% 
  dplyr::mutate(hammer_price_bp_usd_range = forcats::fct_relevel(hammer_price_bp_usd_range, "<$50,000"))%>%
  dplyr::mutate(hammer_price_bp_usd_range = forcats::fct_relevel(hammer_price_bp_usd_range, "<$500,000"))%>%
   dplyr::mutate(hammer_price_bp_usd_range = forcats::fct_relevel(hammer_price_bp_usd_range, "$500,000+"))



vcd::mosaic(hammer_price_bp_usd_range~is_untitled, direction = c("v", "h"),df1,
       gp = gpar(fill = c("lightyellow", "lightpink")),
                 labeling = labeling_border(rot_labels = c(0, 90),pos_labels="center"))

```
#### Are they mostly contemprary? 
```{r}
#untitle_labels = read.csv("untitle_labels", header=TRUE)
```
### Does the era of the lot affect its price?
```{r}
df3 <- df2 %>% 
  filter(df2$birth_year>1800)
ggplot(df3, aes(birth_year,hammer_price_bp_usd))  + 
  geom_smooth(method='lm',formula=y ~ poly(x, 2))+
  geom_point(alpha = .1)  + 
  theme_grey(10)+scale_y_log10()+  geom_density_2d(bins = 5)
```
```{r}
#box plot price vs year
#ggplot(df3, aes(x=auth_era), y=hammer_price_bp_usd)+ stat_summary(fun.y="mean", geom="line", aes(group=1))
ggplot(df3, aes(auth_era, hammer_price_bp_usd)) +
  geom_boxplot()+theme(axis.text.x = element_text(angle = 90, hjust = 1))+scale_y_log10()
```
## Do Certain Extrernal Factors Result in Higher Price?
### Is there an impact from the financial crisis?
```{r}
df1$auc_ymd <- as.Date(df1$auc_year_month_date)
art_yearfin <- df1 %>% group_by(month=floor_date(auc_ymd, "month")) %>% summarise(revenue = mean(hammer_price_bp_usd))
ggplot() +geom_line(data=art_yearfin, aes(x=month, y=revenue/1000000))+ggtitle("Financial Crisis' Effect on average lot price ($M)")+ylab("average price")+xlab("Time")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+scale_y_log10()

```
###Does the Order Matter?
```{r}
ggplot(df2, aes(percent_in_auction,hammer_price_bp_usd))  + 
  geom_smooth(method='lm',formula=y~x,color="red")+
  geom_point(alpha = .05)  + 
  theme_grey(10)+scale_y_log10()+
  facet_wrap(~location)
```
###Does Season Matter?
```{r}

vcd::mosaic(hammer_price_bp_usd_range~auc_season, direction = c("v", "h"),df1,
       gp = gpar(fill = c("lightyellow", "lightpink")),
                 labeling = labeling_border(rot_labels = c(0, 90),pos_labels="center"))
```
###Does Location Matter?
```{r}
```
### How Accurate are the Estimates?
